Arc Blanc: a real time ocean simulation framework

### 2025-01-10 paper.tex
Fixed case in equation 56
Make the order of parameters more consistent k,x < y < t

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Water height}
\label{subsec:waterHeight}
Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999} proposed to solve Equation~\ref{equa:systemTessendorf} using a sum of waves of different amplitudes and wavelengths as follows:
\begin{equation}
	\label{equa:heightWater}
	h(\mathbf x,t)=\sum_{\mathbf{k}} \tilde{h}(t, \mathbf{k}) \exp\left( i \mathbf{k}\cdot \mathbf x\right),
\end{equation}
where $\mathbf{k}=\left[k_x\ k_z\right]^T$ is the wave vector with $k_x=\frac{2\pi n}{L_{i}}$ and $k_z=\frac{2\pi m}{L_{i}}$, $L_i$ is the length of the i-th cascade as defined in Section~\ref{subsec:cascades}, $n$ and $m$ are integers in $\left[-\frac{N}{2},\frac{N}{2}\right]$, $N\in\mathbb N^+$ being a constant given by user, and $\tilde{h}$ is given by:
\begin{equation}
	\label{equa:heightWaterBis}
	\tilde{h}(t, \mathbf{k}) = \tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)+\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t),
\end{equation}
where $k$ is the Euclidean norm of the wave vector $\mathbf{k}$, $\omega(k)=\sqrt{gk}$ is the \textbf{dispersion relation}\footnote{The dispersion relation is defined for deep water, see Appendix~\ref{sec:annexe} for more details.} and $\tilde{h}_0^*$ the conjugate of $\tilde{h}_0$.
Researchers in oceanography have made statistical measurements on the ocean to compute $\tilde{h}_0$, leading to many different \textit{ocean spectra} \cite{tessendorfGilliganPrototypeFramework2017}.
In its original paper, Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999} used the \textit{Phillips Spectrum} with Gaussian fluctuation, leading to the following expression of $\tilde{h}_0(\mathbf{k})$:
\begin{equation}
	\tilde{h}_0(\mathbf{k}) = \xi_{\mathbf{k}}\sqrt{\frac{4\pi}{L_ik}S(\omega)D(\omega, \theta)\left|\frac{\partial\omega(k)}{\partial k}\right|}
\end{equation}
where $\xi_{\mathbf{k}}$ are samples from a Gaussian distribution with mean $0$ and standard deviation $1$, $S(\omega)$ and $D(\omega, \theta)$ are the frequency and directional spectra described respectively in paragraphes \ref{subsubsec:jonswap} and \ref{subsubsec:donelanbanner}.

@@

\Framework uses the horizontal displacement $\mathbf{D}(\mathbf x,t) = \left[D_x\ D_z\right]^T$ of seawater to simulate more realistic choppy waves.
As proposed in Sections 4.5 and 4.6 of Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999}, it is calculated as follows:
\begin{equation}\label{equa:hori-displacement}
	\mathbf{D}(\mathbf x,t) = \sum_{\mathbf{k}} \frac{i\mathbf{k}}{k} \tilde{h}(t, \mathbf{k})\exp\left(i\mathbf{k}\cdot\mathbf x\right).
\end{equation}

@@

\begin{align*}
	\frac{\partial D_x}{\partial x}(\mathbf x,t) & = \sum_{\mathbf{k}} -\frac{k_x^2}{k} \tilde{h}(t, \mathbf{k})\exp\left( i \mathbf{k}\cdot \mathbf x\right),  \\
	\frac{\partial D_z}{\partial x}(\mathbf x,t) & = \sum_{\mathbf{k}} -\frac{k_xk_z}{k} \tilde{h}(t, \mathbf{k})\exp\left( i \mathbf{k}\cdot \mathbf x\right), \\
	\frac{\partial D_z}{\partial z}(\mathbf x,t) & = \sum_{\mathbf{k}} -\frac{k_z^2}{k} \tilde{h}(t, \mathbf{k})\exp\left( i \mathbf{k}\cdot \mathbf x\right),  \\
	\frac{\partial h}{\partial x}(\mathbf x,t)   & = \sum_{\mathbf{k}} ik_x \tilde{h}(t, \mathbf{k})\exp\left( i \mathbf{k}\cdot \mathbf x\right),              \\
	\frac{\partial h}{\partial z}(\mathbf x,t)   & = \sum_{\mathbf{k}} ik_z \tilde{h}(t, \mathbf{k})\exp\left( i \mathbf{k}\cdot \mathbf x\right).
\end{align*}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Velocity of water}
\label{subsec:velWater}
Starting from the system \ref{equa:systemTessendorf}, \framework proposes to compute the velocity of the water at any depth $y$, as follows\footnote{See Appendix   \ref{sec:annexe} for more details on the origin of this expression}:
\begin{equation}\label{equa:velocity}
	% \left[
	%   \begin{array}{c} \displaystyle v_x(\mathbf x,y,t) \\ \displaystyle v_y(x,y,z,t) \\ \displaystyle v_z(x,y,z,t) \end{array}
	% \right]
	\mathbf v(\mathbf x,y,t)=
	\sum_{\mathbf{k}} \tilde{\mathbf{v}}(\mathbf{k},t,y)\exp\left(i\mathbf{k}\cdot\mathbf x\right),
\end{equation}
where the vector $\tilde{\mathbf{v}}(\mathbf{k},t, y)$ is given as follows:
\begin{equation}
	\tilde{\mathbf{v}}\left(\mathbf{k},t,y\right)=%
	E\left(\mathbf{k},y\right)\left(\tilde{h}_0\left(\mathbf{k}\right)\exp\left({i\omega(k)t}\right)-\tilde{h}_0^*\left(-\mathbf{k}\right)\exp\left({-i\omega(k)t}\right)\right)
	\left[
		\begin{array}{c}
			\displaystyle\frac{-k_xg}{\omega(k)} \\
			\displaystyle i\omega(k)             \\
			\displaystyle\frac{-k_zg}{\omega(k)}
		\end{array}
		\right],
\end{equation}
where the attenuation function $E(\mathbf{k},y)\in\mathbb R$ is defined as follows:
\begin{equation}
	\label{equa:attenuationDeep}
	E(\mathbf{k}, y)=
	\begin{cases}
		1+ky     & \text{if }y>0, \\
		\exp(ky) & \text{else}.
	\end{cases}
\end{equation}
The first case ($y>0$) comes from an extrapolation above the sea surface of the normal case ($y\leq0$).

@@

	\phi(\mathbf{x},y,t) = \tilde{\phi}(y, t,\mathbf{k})\exp(i \mathbf{k}\cdot \mathbf{x}).

@@

The function $\tilde{\phi}(y, t,\mathbf{k})$ is defined as follows:

@@

	\tilde{\phi}(y, t,\mathbf{k})=\frac{ig}{\omega(k)} E(\mathbf{k},y) \left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)-\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right)

@@

	                & = \frac{\cosh(ky)\cosh(kH) + \sinh(kY)\sinh(kH)}{\cosh(kH)}\nonumber \\

@@

Equation~\ref{equa:attenuationDeep1} and the definition of potential velocity $\nabla\phi=\mathbf{v}$ lead to the velocity $\mathbf{v}$. First, function $\tilde{\phi}$ is rewritten as follows:
\begin{equation}
	\tilde{\phi}(y,t,\mathbf{k})=\frac{ig}{\omega(k)} E(\mathbf{k},y) \left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)-\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right).
\end{equation}
Then the gradient of $\phi$ is as follows:
\begin{equation}
	\nabla \phi = \begin{pmatrix}
		v_x \\
		v_y \\
		v_z
	\end{pmatrix} =\tilde{\mathbf{v}}(y,\mathbf{k},t)\exp\left(i\mathbf{k}\cdot\mathbf{x}\right)
\end{equation}
where $\tilde{\mathbf{v}}(\mathbf{k},t,y)$ is the vector equal to:
\begin{equation}
	\tilde{\mathbf{v}}\left(\mathbf{k},t,y\right)=%
	E\left(\mathbf{k},y\right)\left(\tilde{h}_0\left(\mathbf{k}\right)\exp\left({i\omega(k)t}\right)-\tilde{h}_0^*\left(-\mathbf{k}\right)\exp\left({-i\omega(k)t}\right)\right)
	\left[
		\begin{array}{c}
			\displaystyle\frac{-k_xg}{\omega(k)} \\
			\displaystyle i\omega(k)             \\
			\displaystyle\frac{-k_zg}{\omega(k)}
		\end{array}
		\right],
\end{equation}

Finally, reintroducing a sum on a set of waves vectors $\mathbf{k}$ leads to:
\begin{equation}
	\nabla \phi =
	\sum_{\mathbf{k}}\tilde{\mathbf{v}}(\mathbf{k},t, y)\exp\left(i\mathbf{k}\cdot\mathbf{x}\right).
\end{equation}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Water height}
\label{subsec:waterHeight}
Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999} proposed to solve Equation~\ref{equa:systemTessendorf} using a sum of waves of different amplitudes and wavelengths as follows:
\begin{equation}
	\label{equa:heightWater}
	h(\mathbf x,t)=\sum_{\mathbf{k}} \tilde{h}(\mathbf{k}, t) \exp\left( i \mathbf{k}\cdot \mathbf x\right),
\end{equation}
where $\mathbf{k}=\left[k_x\ k_z\right]^T$ is the wave vector with $k_x=\frac{2\pi n}{L_{i}}$ and $k_z=\frac{2\pi m}{L_{i}}$, $L_i$ is the length of the i-th cascade as defined in Section~\ref{subsec:cascades}, $n$ and $m$ are integers in $\left[-\frac{N}{2},\frac{N}{2}\right]$, $N\in\mathbb N^+$ being a constant given by user, and $\tilde{h}$ is given by:
\begin{equation}
	\label{equa:heightWaterBis}
	\tilde{h}(\mathbf{k}, t) = \tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)+\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t),
\end{equation}
where $k$ is the Euclidean norm of the wave vector $\mathbf{k}$, $\omega(k)=\sqrt{gk}$ is the \textbf{dispersion relation}\footnote{The dispersion relation is defined for deep water, see Appendix~\ref{sec:annexe} for more details.} and $\tilde{h}_0^*$ the conjugate of $\tilde{h}_0$.
Researchers in oceanography have made statistical measurements on the ocean to compute $\tilde{h}_0$, leading to many different \textit{ocean spectra} \cite{tessendorfGilliganPrototypeFramework2017}.
In its original paper, Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999} used the \textit{Phillips Spectrum} with Gaussian fluctuation, leading to the following expression of $\tilde{h}_0(\mathbf{k})$:
\begin{equation}
	\tilde{h}_0(\mathbf{k}) = \xi_{\mathbf{k}}\sqrt{\frac{4\pi}{L_ik}S(\omega)D(\omega, \theta)\left|\frac{\partial\omega(k)}{\partial k}\right|}
\end{equation}
where $\xi_{\mathbf{k}}$ are samples from a Gaussian distribution with mean $0$ and standard deviation $1$, $S(\omega)$ and $D(\omega, \theta)$ are the frequency and directional spectra described respectively in paragraphes \ref{subsubsec:jonswap} and \ref{subsubsec:donelanbanner}.

@@

\Framework uses the horizontal displacement $\mathbf{D}(\mathbf x,t) = \left[D_x\ D_z\right]^T$ of seawater to simulate more realistic choppy waves.
As proposed in Sections 4.5 and 4.6 of Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999}, it is calculated as follows:
\begin{equation}\label{equa:hori-displacement}
	\mathbf{D}(\mathbf x,t) = \sum_{\mathbf{k}} \frac{i\mathbf{k}}{k} \tilde{h}(\mathbf{k}, t)\exp\left(i\mathbf{k}\cdot\mathbf x\right).
\end{equation}

@@

\begin{align*}
	\frac{\partial D_x}{\partial x}(\mathbf x,t) & = \sum_{\mathbf{k}} -\frac{k_x^2}{k} \tilde{h}(\mathbf{k}, t)\exp\left( i \mathbf{k}\cdot \mathbf x\right),  \\
	\frac{\partial D_z}{\partial x}(\mathbf x,t) & = \sum_{\mathbf{k}} -\frac{k_xk_z}{k} \tilde{h}(\mathbf{k}, t)\exp\left( i \mathbf{k}\cdot \mathbf x\right), \\
	\frac{\partial D_z}{\partial z}(\mathbf x,t) & = \sum_{\mathbf{k}} -\frac{k_z^2}{k} \tilde{h}(\mathbf{k}, t)\exp\left( i \mathbf{k}\cdot \mathbf x\right),  \\
	\frac{\partial h}{\partial x}(\mathbf x,t)   & = \sum_{\mathbf{k}} ik_x \tilde{h}(\mathbf{k}, t)\exp\left( i \mathbf{k}\cdot \mathbf x\right),              \\
	\frac{\partial h}{\partial z}(\mathbf x,t)   & = \sum_{\mathbf{k}} ik_z \tilde{h}(\mathbf{k}, t)\exp\left( i \mathbf{k}\cdot \mathbf x\right).
\end{align*}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Velocity of water}
\label{subsec:velWater}
Starting from the system \ref{equa:systemTessendorf}, \framework proposes to compute the velocity of the water at any depth $y$, as follows\footnote{See Appendix   \ref{sec:annexe} for more details on the origin of this expression}:
\begin{equation}\label{equa:velocity}
	% \left[
	%   \begin{array}{c} \displaystyle v_x(\mathbf x,y,t) \\ \displaystyle v_y(x,y,z,t) \\ \displaystyle v_z(x,y,z,t) \end{array}
	% \right]
	\mathbf v(\mathbf x,y,t)=
	\sum_{\mathbf{k}} \tilde{\mathbf{v}}(\mathbf{k}, y, t)\exp\left(i\mathbf{k}\cdot\mathbf x\right),
\end{equation}
where the vector $\tilde{\mathbf{v}}(\mathbf{k}, y, t)$ is given as follows:
\begin{equation}
	\tilde{\mathbf{v}}\left(\mathbf{k}, y, t\right)=%
	E\left(\mathbf{k},y\right)\left(\tilde{h}_0\left(\mathbf{k}\right)\exp\left({i\omega(k)t}\right)-\tilde{h}_0^*\left(-\mathbf{k}\right)\exp\left({-i\omega(k)t}\right)\right)
	\left[
		\begin{array}{c}
			\displaystyle\frac{-k_xg}{\omega(k)} \\
			\displaystyle i\omega(k)             \\
			\displaystyle\frac{-k_zg}{\omega(k)}
		\end{array}
		\right],
\end{equation}
where the attenuation function $E(\mathbf{k},y)\in\mathbb R$ is defined as follows:
\begin{equation}
	\label{equa:attenuationDeep}
	E(\mathbf{k}, y)=
	\begin{cases}
		1+ky     & \text{if }y>0, \\
		\exp(ky) & \text{else}.
	\end{cases}
\end{equation}
The first case ($y>0$) comes from an extrapolation above the sea surface of the normal case ($y\leq0$).

@@

	\phi(\mathbf{x},y,t) = \tilde{\phi}(\mathbf{k}, y, t)\exp(i \mathbf{k}\cdot \mathbf{x}).

@@

The function $\tilde{\phi}(\mathbf{k}, y, t)$ is defined as follows:

@@

	\tilde{\phi}(\mathbf{k}, y, t)=\frac{ig}{\omega(k)} E(\mathbf{k},y) \left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)-\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right)

@@

	                & = \frac{\cosh(ky)\cosh(kH) + \sinh(ky)\sinh(kH)}{\cosh(kH)}\nonumber \\

@@

Equation~\ref{equa:attenuationDeep1} and the definition of potential velocity $\nabla\phi=\mathbf{v}$ lead to the velocity $\mathbf{v}$. First, function $\tilde{\phi}$ is rewritten as follows:
\begin{equation}
	\tilde{\phi}(\mathbf{k}, y, t)=\frac{ig}{\omega(k)} E(\mathbf{k},y) \left(\tilde{h}_0(\mathbf{k})\exp(i\omega(k)t)-\tilde{h}_0^*(-\mathbf{k})\exp(-i\omega(k)t)\right).
\end{equation}
Then the gradient of $\phi$ is as follows:
\begin{equation}
	\nabla \phi = \begin{pmatrix}
		v_x \\
		v_y \\
		v_z
	\end{pmatrix} =\tilde{\mathbf{v}}(\mathbf{k}, y, t)\exp\left(i\mathbf{k}\cdot\mathbf{x}\right)
\end{equation}
where $\tilde{\mathbf{v}}(\mathbf{k}, y, t)$ is the vector equal to:
\begin{equation}
	\tilde{\mathbf{v}}\left(\mathbf{k}, y, t\right)=%
	E\left(\mathbf{k},y\right)\left(\tilde{h}_0\left(\mathbf{k}\right)\exp\left({i\omega(k)t}\right)-\tilde{h}_0^*\left(-\mathbf{k}\right)\exp\left({-i\omega(k)t}\right)\right)
	\left[
		\begin{array}{c}
			\displaystyle\frac{-k_xg}{\omega(k)} \\
			\displaystyle i\omega(k)             \\
			\displaystyle\frac{-k_zg}{\omega(k)}
		\end{array}
		\right],
\end{equation}

Finally, reintroducing a sum on a set of waves vectors $\mathbf{k}$ leads to:
\begin{equation}
	\nabla \phi =
	\sum_{\mathbf{k}}\tilde{\mathbf{v}}(\mathbf{k}, y, t)\exp\left(i\mathbf{k}\cdot\mathbf{x}\right).
\end{equation}
